---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Fast Bayesian Inference for Change-Point Detection

<!-- badges: start -->
<!-- badges: end -->

`{mich}` is an R package that implements the Multiple Independent Change-Point 
(MICH) method introduced in [Berlind, Cappello, Madrid Padilla (2025)](https://arxiv.org/abs/2507.01558). 
The main function in the package `mich()` takes a length $T$ sequence of 
observations $\mathbf{y}_{1:T}$ with potentially many change-points in the mean 
and variance, and deploys a backfitting procedure to find a variational 
approximation to the posterior distribution of the change-points.

## Installation

You can install the development version of `mich` from GitHub using:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("davis-berlind/mich")
```


