% document set up
\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\setlength{\parindent}{0cm}
\setlength{\parskip}{1em}

% math packages
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumitem}
\usepackage{bbm}

% general packages
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{natbib}

% algorithms
\usepackage[ruled,vlined]{algorithm2e}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
\SetCommentSty{mycommfont}


% plate diagram
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\newcommand\indep{\perp\!\!\!\perp}
\newcommand{\sforall}{\;\forall\;}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\Pr}{\mathbb{P}}
\newcommand{\Var}{\text{\normalfont Var}}
\newcommand{\Cov}{\text{\normalfont Cov}}
\newcommand{\argmin}[1]{\underset{#1}{\text{\normalfont arg min}}}
\newcommand{\argmax}[1]{\underset{#1}{\text{\normalfont arg max}}}

\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

% Title
\usepackage{authblk}

\title{Variational Bayesian Change-Point Detection with Interpretable Credible Sets in Fixed- and High-Dimension}

\author[1]{Davis Berlind}
\author[2]{Lorenzo Cappello}
\author[3]{Oscar Hernan Madrid Padilla}
\affil[1]{Department of Statistics, University of California Los Angeles}
\affil[2]{Department of Economics and Business, Universitat Pompeu Fabra; Data Science Center, Barcelona School of Economics}
\affil[3]{Department of Statistics, University of California Los Angeles}


\date{\today}

\begin{document}

\numberwithin{equation}{section}

\maketitle

\begin{abstract}
    We propose a novel variational Bayes procedure for detecting structural breaks in the mean and variance of a sub-Gaussian sequence. Our method returns credible sets around the estimated locations of the change-points, enabling interpretable uncertainty quantification. We introduce an efficient back-fitting algorithm for approximating the posterior distribution of the change-point locations that is an order of magnitude faster than existing Bayesian methods. We also study the asymptotic behavior of our method in the special case of a single change-point. Under standard regularity conditions, we prove that maximum \textit{a posteriori} estimator of the change-point locations returned by our model is consistent in both the fixed- and high-dimensional settings when there is a change in the mean, and in the univariate setting when there is a change in the variance. In each case we show that the localization rate achieved by our method is minimax optimal up to a logarithmic factor. Extensive simulation studies demonstrate that our method is competitive with the state-of-the-art and in many cases it provides the narrowest credible sets without sacrificing nominal coverage guarantees. We test the method in a practical setting by applying it to the problem of detecting structural changes in the lithology of an oil well. 
\end{abstract}

\input{MICH/1. Introduction}

\input{MICH/2. Single Change-Point Models}

\input{MICH/3. MICH}

\input{MICH/4. Simulations}

\input{MICH/5. Data}

\input{MICH/6. Discussion}

\bibliographystyle{chicago}

\bibliography{MICH/references}

\appendix

\input{MICH/7. Appendix.tex}

\end{document}