---
title: "Detecting Changes in Daily Step Count"
author: "Davis Berlind"
date: "`r Sys.Date()`"
description: "Using MICH to detect changes in my daily step count."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detecting Changes in Daily Step Count}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
---
title: "Detecting Changes in Daily Step Count"
author: "Davis Berlind"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detecting Changes in Daily Step Count}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Due to the widespread adoption of wearable devices over the past decade, individuals now have unprecedented access to high-frequency personal health data. A [2023 NIH survey](https://www.nhlbi.nih.gov/news/2023/study-reveals-wearable-device-trends-among-us-adults) revealed that "one in three Americans uses a wearable device, such as a smart watch or band, to track their health and fitness." The data collected from these devices can be used to detect changes in a person's health or lifestyle. I don't personally own any wearable devices, but for nearly a decade I have used my phone to record the steps I take in a given day. This data is bundled with the `{mich}` package as the `daily_steps` data set. Below I show how I used `mich` to detect the changes in my life that manifest in `daily_steps`. In [Berlind, Cappello, and Madrid Padilla (2025)](https://arxiv.org/abs/2507.01558), we showed how to extend the MICH model to count data, but seeing as this feature is not yet implemented in the `{mich}` package, I instead model the change-points as joint changes to the mean and variance of `daily_steps$total_steps` and set `J_auto == TRUE` to automatically select the number of mean-variance components to include in the model. 

```{r fit-model, cache = TRUE}
library(mich)
daily_steps$day <- as.Date(daily_steps$day)

step_fit <- mich(daily_steps$total_steps, J_auto = TRUE, tol = 1e-12,
                 max_iter = Inf, restart = FALSE)
```
```{r plot-model, echo = FALSE, dpi=500, out.width="99%",}
par(oma = c(0,0,0,0), mar = c(2,2,2,0.5), tcl = -0.2, mgp = c(1.5, 0.1, 0))
cred_sets <- mich_sets(step_fit$meanvar_model$pi_bar, level = 0.9, max_length = log(nrow(daily_steps))^2)
est_cp <- cred_sets$cp
plot(daily_steps$day, daily_steps$total_steps, type = "l", 
     main = "Daily Steps Sept. 2016 - Aug. 2025", 
     xlab = "", ylab = "", col = "white", cex.main=0.5, 
     cex.lab=0.7, cex.axis=0.5, xaxt = "n")
for(i in unlist(cred_sets$sets)) { 
  if (i == 1) next
  rect(xleft = daily_steps$day[i-1], xright = daily_steps$day[i+1], 
       ybottom = par("usr")[3], ytop = par("usr")[4], 
       col =  adjustcolor("lightblue", alpha = 1), 
       border = NA)
}
lines(daily_steps$day, daily_steps$total_steps, type = "l", lwd = 0.5)
lines(daily_steps$day, step_fit$mu, lwd = 1, col = "red")
abline(v = daily_steps$day[est_cp], lwd = 1, lty = 2, col = "blue")
axis(1, at = seq(min(daily_steps$day), max(daily_steps$day), by = "quarters") , 
     labels = FALSE)
text(
  x = seq(min(daily_steps$day), max(daily_steps$day), by = "quarters"), 
  y = par("usr")[3] - 1000,    # Position just below axis
  labels = format(seq(min(daily_steps$day), max(daily_steps$day), by = "quarters"), "%b %Y"),
  srt = 45,                            # Rotation in degrees
  xpd = TRUE,                          # Allow drawing outside plot region
  adj = 1.1,                              # Adjust alignment
  cex = 0.3
)
```

After fitting `mich()` to the step data, we see that there are 12 changes detected, each of which is annotated in the table below:

| Estimated Change | 90\% CS Size | Annotation |
|:---|--:|:-----------|
| `r format(daily_steps$day[est_cp][1], "%B %d, %Y")` | `r length(cred_sets$sets[[1]])` | I lived in NYC from 2016 to 2018 and walked much more during the summer. This change-point appears to be this seasonal change (note the wide 90\% credible set that extends from March to June). |
| `r format(daily_steps$day[est_cp][2], "%B %d, %Y")` | `r length(cred_sets$sets[[2]])` | Again, this change seems to be the switch from winter to summer (note that MICH apparently misses the seasonal decrease in the winter of 2017).|
| `r format(daily_steps$day[est_cp][3], "%B %d, %Y")` | `r length(cred_sets$sets[[3]])` | I moved from NYC to Durham, NC at the beginning of August 2018, to begin a masters degree program at Duke (orientation was August 21, 2018).|
| `r format(daily_steps$day[est_cp][3], "%B %d, %Y")` | `r length(cred_sets$sets[[3]])` | I moved from NYC to Durham, NC at the beginning of August 2018, to begin a masters degree program at Duke (orientation was August 21, 2018).|
| `r format(daily_steps$day[est_cp][4], "%B %d, %Y")` | `r length(cred_sets$sets[[4]])` | From 8-12-2019 to 8-24-2019 I traveled to Ireland and Scotland with my now wife.|
| `r format(daily_steps$day[est_cp][5], "%B %d, %Y")` | `r length(cred_sets$sets[[5]])` | Return from Ireland/Scotland. |
| `r format(daily_steps$day[est_cp][6], "%B %d, %Y")` | `r length(cred_sets$sets[[6]])` | On 3-17-2020 I flew to Miami to spend the first phase of lock-down with my wife and her family. |
| `r format(daily_steps$day[est_cp][7], "%B %d, %Y")` | `r length(cred_sets$sets[[7]])` | On 6-11-2020 I moved to Boston. |
| `r format(daily_steps$day[est_cp][8], "%B %d, %Y")` | `r length(cred_sets$sets[[8]])` | On 9-2-2020 I began a PhD in economics at BU. |
| `r format(daily_steps$day[est_cp][9], "%B %d, %Y")` | `r length(cred_sets$sets[[9]])` | On 5-8-2021 I took my last exam at BU and left the program to transfer to the Department of Statistics at UCLA. |
| `r format(daily_steps$day[est_cp][10], "%B %d, %Y")` | `r length(cred_sets$sets[[10]])` | From 6-29-2023 to 7-13-2023 I traveled to Prague, Vienna, and Greece for a friend's wedding. |
| `r format(daily_steps$day[est_cp][11], "%B %d, %Y")` | `r length(cred_sets$sets[[11]])` | Returned from Greece. |
| `r format(daily_steps$day[est_cp][12], "%B %d, %Y")` | `r length(cred_sets$sets[[12]])` | This change captures the fact that in May I hunkered down and barely left my desk in order to get the preprint of my paper up on arXiv 💀. |
